Building socket-aware BPF programs

### 2018-10-27 paper/paper.tex
paper: Add stateful firewalling

    This paper presents the changes necessary to extend socket introspection
    capabilities of the Linux kernel's BPF packet hooks. To provide safety
    guarantees around the access of the socket, the verifier is extended to
    track allocation and release of kernel resources. Using this mechanism, we
    introduce socket lookup and release helpers into the BPF API.

    %These new primitives open up new possibilities for various use cases
    %including stateful firewalling and load balancing.

    %To provide access to socket state from a BPF helper call, the verifier
    %needs the ability to track the safety of the calls, including appropriate
    %reference counting upon the underlying socket for the duration of the
    %access.

    %With such an ability ability in the verifier, we add socket lookup and
    %release function calls to the BPF API, allowing BPF programs to safely find
    %a socket and build logic upon the presence or attributes of a socket.
    %Applications for this functionality range from load-balancing traffic based
    %on the presence of a listening application to building stateful firewalling
    %primitives to understand whether traffic for this connection has been seen
    %before. With this new functionality, BPF programs can integrate more
    %closely with the networking stack's understanding of the traffic transiting
    %the kernel.

    %provide safe access to sockets from BPF packet handling hooks. The BPF verifier
    %is extended to gain understanding of reference-counted pointer types, which
    %allows BPF code paths to

@@

\begin{itemize}
\item BPF started with sockets, but only provided access to packets
\item Previous helper extensions allowed e.g. socket operation handling
\item Arbitrary socket lookup from socket/packet lookups allows more powerful
      BPF programs
\end{itemize}

\todo{Rewrite introduction}

% Over the past several years, BPF has steadily become more powerful in
% multiple ways: Through building more intelligence into the verifier which
% allows more complex programs to be loaded, and through extension of the API
% such as by adding new map types and new native BPF function calls. While
% BPF has its roots in applying filters at the socket layer, the ability to
% introspect the sockets relating to traffic being filtered has been limited.

%To build such awareness into a BPF helper, the verifier needs the ability to
%track the safety of the calls, including appropriate access into socket
%pointer types, and reference counting upon the underlying socket. This paper
%walks through extensions to the verifier to perform tracking of references in
%a BPF program. This allows BPF developers to extend the API with functions
%that allocate and release resources within the execution lifetime of a BPF
%program, and the verifier will validate that the resource acquisition and
%release is balanced across all execution paths of the BPF programs.

%Using this new reference tracking ability in the verifier, we add socket lookup
%and release function calls to the BPF API, allowing BPF programs to safely find
%a socket and build logic upon the presence or attributes of a socket. This can
%be used to load-balance traffic based on the presence of a listening
%application, or to implement stateful firewalling primitives to understand
%whether traffic for this connection has been seen before. With this new
%functionality, BPF programs can integrate more closely with the networking
%stack's understanding of the traffic transiting the kernel.

\todo{Desire to target logic based upon socket existence/attributes at packet
layers}

@@

The BPF verifier must ensure that all BPF programs that are loaded into the
kernel are safe to execute. For accessing socket pointers, this requires
validation that memory accesses are within the bounds of the memory allocated
for the structure, and will not make any modifications to the socket fields
that would be incompatible with the core socket handling logic. This is
provided through extension of existing functionality which provides bounds
verification and offset conversion for programs operating on CGroups hooks.
Additionally, the underlying socket memory must remain associated with the
socket for the duration of accesses from the BPF program, which may be
guaranteed by taking a reference on sockets for the duration of their use.

@@

As with many other structure types, Linux sockets are reference-counted to
track usage and handle memory release. In general to ensure the safety of
socket for the duration of access in BPF programs, this reference count needs
to be incremented when sockets are referenced from BPF programs. Two options
were considered for how to guarantee this: Implicit reference tracking and
explicit reference tracking.

@@

>>>>

    To build such awareness into a BPF helper, the verifier needs the ability
    to track the safety of the calls, including appropriate reference counting
    upon the underlying socket. This paper describes extensions to the verifier
    to perform tracking of references in a BPF program. This allows BPF
    developers to extend the UAPI with functions that allocate and release
    resources within the execution lifetime of a BPF program, and the verifier
    will validate that the resources are released exactly once prior to program
    completion.

    Using this new reference tracking ability in the verifier, we add socket
    lookup and release function calls to the BPF API, allowing BPF programs to
    safely find a socket and build logic upon the presence or attributes of a
    socket. This can be used to load-balance traffic based on the presence of a
    listening application, or to implement stateful firewalling primitives to
    understand whether traffic for this connection has been seen before. With
    this new functionality, BPF programs can integrate more closely with the
    networking stack's understanding of the traffic transiting the kernel.

@@

\begin{itemize}
\item BPF started with sockets, but only provided access to packets
\item Previous helper extensions allowed e.g. socket operation handling
\item Arbitrary socket lookup from socket/packet hooks allows more powerful BPF
      programs
\item Use cases revolve around building logic in socket/packet layer hooks that
      change the behaviour based on the existence and attributes of a socket.
\end{itemize}

@@

The BPF verifier must ensure that all BPF programs that are loaded into the
kernel are safe to execute. For accessing socket pointers, this requires
validation that memory accesses are within the bounds of the memory allocated
for the structure, and will not make any modifications to the socket fields
that would be incompatible with the core socket handling logic. This is
provided through extension of existing functionality which performs bounds
verification and offset conversion for programs operating on CGroups hooks.
Additionally, the underlying memory must remain associated with the socket for
the duration of accesses from the BPF program, which may be guaranteed by
taking a reference on sockets for the duration of their use. The following
sections explain how this is implemented through the introduction of a new
pointer type in the verifier, and through the extension of the verifier to
recognise assembly instructions that represent reference acquisition and
release.

@@

Sockets in Linux are reference-counted to track usage and handle memory
release. In general to ensure the safety of socket for the duration of access
in BPF programs, this reference count needs to be incremented when sockets are
referenced from BPF programs. Two options were considered for how to guarantee
this: Implicit reference tracking and explicit reference tracking.

@@

Stateful firewalling is distinguished from stateless firewalling in that it
provides the ability to associate two directions of a connection together and
apply a filtering policy based upon the direction of the connection. This
commonly involves using a connection tracker to reconstruct the connection
state of each endpoint of the connection without specific knowledge from those
endpoints~\cite{ovs-ct,conntrack}. In cases where the Linux stack acts as a
firewall on a node in the network that is separate from both endpoints, In many
cases this model is used when filtering traffic where one endpoint is colocated
with the Linux stack that performs the firewalling, which already contains
information about the local end of the connection in a slightly.
